---
title: "DATA 607 SQL and R: Movie Ratings"
author: "Zineb Tamnat"
output: html_document
---

## Overview

For this assignment I collected 1–5 movie ratings from several people, stored them in a PostgreSQL database, and analyzed them in R. Because not everyone has seen every movie, missing ratings are expected and handled as NA in R.

## Load Packages
```{r}
library(DBI)
library(RPostgres)
library(dplyr)
library(tidyr)
```

I connected to my database.

```{r}
con <- dbConnect(
  RPostgres::Postgres(),
  host = Sys.getenv("PGHOST"),
  port = Sys.getenv("PGPORT"),
  dbname = Sys.getenv("PGDATABASE"),
  user = Sys.getenv("PGUSER"),
  password = Sys.getenv("PGPASSWORD")
)
dbListTables(con)
```

Here I am pulling the ratings from the SQL.
```{r}
ratings_long <- dbGetQuery(con, "
SELECT u.name, m.title, r.rating
FROM ratings r
JOIN users u  ON u.user_id = r.user_id
JOIN movies m ON m.movie_id = r.movie_id
ORDER BY u.name, m.title;
")

ratings_long
```


Here I am creating the full User and Movie Table while handling  missing ratings.
```{r}
users  <- dbGetQuery(con, "SELECT name FROM users ORDER BY name;")
movies <- dbGetQuery(con, "SELECT title FROM movies ORDER BY title;")

ratings_complete <- expand_grid(
  name = users$name,
  title = movies$title
) %>%
  left_join(ratings_long, by = c("name", "title"))

ratings_complete
```

Here I am getting the average ratings
```{r}
movie_summary <- ratings_complete %>%
  group_by(title) %>%
  summarise(
    num_ratings = sum(!is.na(rating)),
    avg_rating = mean(rating, na.rm = TRUE)
  ) %>%
  arrange(desc(avg_rating))

movie_summary
```

Here I am getting the ratings per person
```{r}
user_summary <- ratings_complete %>%
  group_by(name) %>%
  summarise(
    num_rated = sum(!is.na(rating)),
    avg_given = mean(rating, na.rm = TRUE)
  )

user_summary
```









